---
title: "Practical Machine Learning Project"
author: "Braydon"
date: "1/12/2020"
output:
  html_document: default
  pdf_document: default
---
# Overview

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

# Prep Work

```{r, echo = TRUE}
rm(list=ls())
library(dplyr)
library(knitr)
library(caret)
library(corrplot)
library(randomForest)
library(rpart)
library(rpart.plot)
library(rattle)
library(gbm)
```

# Load Data & Create Training/Test Set

```{r, echo=TRUE}
# Load the data
UrlTrain <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
UrlTest  <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

training <- read.csv(url(UrlTrain))
testing <- read.csv(url(UrlTest))

# Create data partition

inTrain <- createDataPartition(training$classe,p=0.7,list=FALSE)
train <- training[inTrain,]
test <- training[-inTrain,]
```

# Remove vaiables that are mostly NA and near zero values

```{r,echo=TRUE}
# Near Zero Variance Columns

near_zero <- nearZeroVar(train)
train <- train[,-near_zero]
test <- test[,-near_zero]

# NA Columns with 95% NA values

AllNA <- sapply(train, function(x) mean(is.na(x))) > 0.95
train <- train[,AllNA==FALSE]
test <- test[,AllNA==FALSE]

# Remove PII Columns

train <- train[,-(1:5)]
test <- test[,-(1:5)]
```

# Variable Correlation

```{r,echo=TRUE}
# Correlation prior to moving on to modelling

corMatrix <- cor(train[,-54])
corrplot(corMatrix,order = "FPC",method = "color", type = "lower",tl.cex = 0.8, tl.col = rgb(0,0,0))
```


# Prediction Models

```{r, echo=TRUE}
# Method: Random Forest - Fitting the Model

set.seed(1234)
RFcontrol <- trainControl(method = "cv", number=5,verboseIter = FALSE)
FitRFmodel <- train(classe~., data = train, method = "rf", trControl = RFcontrol)
FitRFmodel$finalModel
plot(FitRFmodel,main="RF Model Accuracy by Predictors")

```

```{r, echo=TRUE}
# Method: Random Forest - Testing the Model
RF_Predict <- predict(FitRFmodel,newdata = test)
RF_cm <- confusionMatrix(test$classe,RF_Predict)
RF_cm
```

## Accuracy is 99.8% which is extremely good. OOB rate is less than 0.2%.

```{r, echo=TRUE}
# Method: Decision Tree - Fitting the Model

DTcontrol <- trainControl(method = "cv", number=5,verboseIter = FALSE)
DT_Model <- train(classe~., data=train, method="rpart", trControl=DTcontrol)
fancyRpartPlot(DT_Model$finalModel)
```

```{r, echo=TRUE}
# Method: Decision Tree - Testing the Model
DT_Predict <- predict(DT_Model,newdata = test)
DT_cm <- confusionMatrix(test$classe,DT_Predict)
DT_cm
```

## The decision tree model produced an accuracy score of 0.52, meaning the out of bag error rate is 0.48. This is not nearly as good as the random forest model ran earlier. 

```{r, echo=TRUE}
# Method: Generalized Boosted Model - Fitting the Model

GBMcontrol <- trainControl(method = "cv", number=3,verboseIter = FALSE)
GBM_Model <- train(classe~., data=train, method="gbm", trControl=GBMcontrol, verbose=FALSE)
plot(GBM_Model)

```

```{r, echo=TRUE}
# Method: Gradient Boosted Model - Testing the Model
GBM_Predict <- predict(GBM_Model,newdata=test)
GBM_cm <- confusionMatrix(test$classe,GBM_Predict)
GBM_cm
```

## Accuracy for the GBM model is 98.5% which is near the Random Forest model above but lies just below by about 1.3%. OOB rate for this model is 1.5%.

# Conclusion 

## By comparing the accuracy rate of all three models, it can be determined that the best fitting model is Random Forest. 
